\thetitle

### 2008-05-29 paper.tex
More work on intro.

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and computers on the Internet and in local area networks~(\caps{LAN}s).
The field of realistic workload generation, however, remains an active area of research with many unanswered questions.
One of the major challenges of modeling full-network traffic is the manifest multilevel interdependence of behaviors.
% The packet is the fundamental unit of traffic in any packet-switched network, including \caps{IP} networks.
% Packets are aggregated at a variety of logical levels: by protocol type, by source and destination nodes, by flows sequences between communicating applications.
% The behaviors of packets are specific to each aggregated group: the distribution of packet sizes is specific to each node as both source and destination, and to each flow as well.
Because of the lack of adequate techniques for modeling this complex multilevel interdependence, three fundamental types of simplifying assumptions are commonly applied to both the analysis and generation of traffic:
\begin{enumerate}
\item \textbf{Uniformity:} that certain aspects of traffic behavior are uniformly distributed. For example, that  all nodes in a network have equal probability of being the source or destination of traffic.
\item \textbf{Independence:} that traffic properties share a single marginal distribution across the entire network, and individual values are independently drawn from this marginal distribution. For example, assuming that a single distribution describes an entire network, as opposed to modeling each flow with its own distribution of packet sizes.
\item \textbf{Subdivided Independence:} that traffic properties behave independently, but only within certain groupings. The assumption, for example, that all flows emanating from each single node share the same distribution of packet sizes.
\end{enumerate}

The rate of packet arrivals at routers on the Internet backbone is known to exhibit complex, self-similar patterns with long-range dependence~\cite{Paxson95,Paxson96,Willinger98}. Other research has demonstrated how to accurately model the behavior of a single flow of packets between two application endpoints, including ``closed-loop'' feedback induced by the ubiquitous transmission control protocol~(\caps{TCP})~\cite{Hernandez06:dissertation}. Recently, the high-level statistical properties of wireless \caps{LAN} traffic have been rigorously modeled with a variety of parametric distributions~\cite{Hernandez06:wlan-traffic}.

>>>>

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and computers on the Internet and in local area networks~(\caps{LAN}s).
Both traffic analysis and realistic workload generation, however, remain active areas of research with many unanswered questions.
One of the major challenges of modeling full-network traffic is the manifest multilevel interdependence of behaviors.
% The packet is the fundamental unit of traffic in any packet-switched network, including \caps{IP} networks.
% Packets are aggregated at a variety of logical levels: by protocol type, by source and destination nodes, by flows sequences between communicating applications.
% The behaviors of packets are specific to each aggregated group: the distribution of packet sizes is specific to each node as both source and destination, and to each flow as well.
% Because of the lack of adequate techniques for modeling this complex multilevel interdependence,
Three types of simplifying assumptions are commonly applied to both analysis and generation of traffic:
\begin{enumerate}
\item \textbf{Uniformity:} that certain aspects of traffic behavior are uniformly distributed. For example, that  all nodes in a network have equal probability of being the source or destination of packets or flows.
\item \textbf{Independence:} that traffic properties share a single marginal distribution across the entire network, and individual values are independently drawn from this marginal distribution. For example, assuming that a single distribution describes an entire network, as opposed to modeling each flow with its own distribution of packet sizes.
\item \textbf{Subdivided Independence:} that traffic properties behave independently, but only within certain groupings. The assumption, for example, that all flows emanating from each single node share the same distribution of packet sizes.
\end{enumerate}
We would clearly be setting up a straw man if we implied that researchers who use these assumptions actually believe them to be true. They clearly are not: flows do not all share a single packet size distribution; nodes are not uniformly likely sources. These assumptions are applied because the drastically simplify both analysis and workload generation. The question remains, however, whether these assumptions are benign. Is the analysis and generation of network traffic drastically distorted by such assumptions? This is the first fundamental question that this paper addresses.

While these assumptions certainly simplify both analysis and workload generation, they are quite certainly not true: flows do not all share a single packet size distribution; every node is not equally likely to be the source or destination of each packet or flow.

% The rate of packet arrivals at routers on the Internet backbone is known to exhibit complex, self-similar patterns with long-range dependence~\cite{Paxson95,Paxson96,Willinger98}. Other research has demonstrated how to accurately model the behavior of a single flow of packets between two application endpoints, including ``closed-loop'' feedback induced by the ubiquitous transmission control protocol~(\caps{TCP})~\cite{Hernandez06:dissertation}. Recently, the high-level statistical properties of wireless \caps{LAN} traffic have been rigorously modeled with a variety of parametric distributions~\cite{Hernandez06:wlan-traffic}.
